\singlespacing{\textbf{Efficient Pollution Abatement in Electricity Markets with Intermittent Renewable Energy}}

### 2019-07-16 documents/draft.tex
Update draft
- update semiparametric notebook robustness checks
- slight changes to results section

	We also consider additional robustness checks for the estimate in fit (3). Firstly, we check if outliers are affecting the results meaningfully. We run the fit (3) regression but trim a larger number of outliers. We find that trimming an additional 1\%, 5\%, or 10\% of outliers does not appear to significantly change the estimate of $\hat{\sigma}$ or meaningfully affect its standard error. Secondly, we test whether any particular states are driving the results. We run regress fit (3) on 48 subsamples; in each subsample, one of the 48 states in our data set is dropped out. We plot the results in the Appendix \autoref{fig:regstaterobust}. The mean and median of the estimates for $\sigma$ are  $0.8873$ and $0.8860$; additionally, 95\% of the estimates lie in $(0.8333, 0.9343)$. All of these estimates are highly significant -- the average $|t|$ value is $19.88$ and the smallest $|t|$ is $18.52$. Furthermore, it seems that most the drop out estimates do not differ significantly from the full sample estimate of $\hat{\sigma} = 0.8847$ (stdev = $0.044$). And, of the subsample estimates that do significantly differ, the magnitude of the difference does not appear to be economically meaningful. To elaborate on this and the practical meaning of $\sigma$, we now discuss what these estimates mean in the context of renewable energy policy.

@@

	We are concerned with how $\sigma$ affects the substitutability of fossil and renewable energy. This is important, because sufficient substitutability between these two technologies is required to transition into greener economy in the future. For instance, \citet{Ace2012} provide a model where they argue that, ``When the two
	sectors [clean and dirty energy] are substitutable but not sufficiently so, preventing an environmental disaster requires a permanent policy intervention. Finally, when the two sectors are complementary, the only way to stave off a disaster is to stop long-run growth." However, note that while our model sets the intertemporal elasticity of substitution for  electricity consumption $\sigma$ to a fixed value, we make no direct assumptions about the elasticity of substitution between different energy technologies. Instead, this latter elasticity is indirectly defined by the substitutability of electricity consumption between periods. To illustrate this, we parametrize and estimate our model numerically.

	Firstly, let technology 1 be coal power and technology 2 be solar power. These two technologies exemplify consistent fossil energy and intermittent renewable energy. Furthermore,  let period $t$ represent peak hours and period $s$ represent off-peak hours. We assume that, holding prices equal, consumers prefer that approximately 60\% of their energy arrive in period $t$ and the remaining 40\% arrive in period $s$; that is, we have $\alpha_t = 0.6$ and $\alpha_s = 0.4$. Next, we normalize all quantity units to a MWh basis. Hence, we let $\xi$ represent the percent of capacity utilized in each period; we assume coal uses 100\% of its capacity in both periods, while solar can access 100\% during peak hours and only 10\% during the off-peak. Finally, we set the cost parameters, given in  \$ per MWh, equal to LCOE\footnote{ The LCOE for a generation technology is equal to the sum of its lifetime costs divided by its lifetime energy output. \citet{Joskow2011} explains the flaws of comparing generation technologies solely on the basis of LCOE; but,  our use of this measure is unrelated to his critique. That is, he argues that the economic value an intermittent technology should also account for when it produces electricity and the prices of electricity in those periods (see Table 2 of his paper). Our model does exactly this; but, we still need to  use LCOE to parametrize the cost of developing capacity. } estimates for 2023 from the \citet{EIALCOE}; specifically, we use estimates for "Solar PV" and "Coal with 30\% CCS" from Table 1b.  In total, we have the following parameters.

>>>>

	We also consider additional robustness checks for the estimate in fit (3). Firstly, we check if outliers are affecting the results meaningfully. We run the fit (3) regression but trim a larger number of outliers. We find that trimming an additional 1\%, 5\%, or 10\% of outliers does not appear to significantly change the estimate of $\hat{\sigma}$ or meaningfully affect its standard error. Secondly, we test whether any particular states are driving the results. We run regress fit (3) on 48 subsamples; in each subsample, one of the 48 states in our data set is dropped out. We plot the results in the Appendix \autoref{fig:regstaterobust}. The mean and median of the estimates for $\sigma$ are  $0.8873$ and $0.8860$; additionally, 95\% of the estimates lie in $(0.8333, 0.9343)$. All of these estimates are highly significant -- the average $|t|$ value is $19.88$ and the smallest $|t|$ is $18.52$. And, although, it seems that some of the estimates are (statistically) significantly  from the full sample estimate of $\hat{\sigma} = 0.8847$ (stdev = $0.044$), the magnitude of the difference does not appear to be economically meaningful. That is, the largest estimate of $\hat{\sigma}$ is 0.9793 which is only 0.0946 larger than the full sample estimate, while the smallest estimate is 0.8174 which is 0.0674 smaller than the full sample estimate. These differences are not meaningful in the practical application of our model. To elaborate on this and the full economic implications of $\sigma$, we now discuss what these estimates mean in the context of renewable energy policy.

@@

	To start, we consider how $\sigma$ affects the substitutability of fossil and renewable energy. This is important, because sufficient substitutability between these two technologies is required to transition into greener economy in the future. For instance, \citet{Ace2012} provide a model where they argue that, ``When the two	sectors [clean and dirty energy] are substitutable but not sufficiently so, preventing an environmental disaster requires a permanent policy intervention. Finally, when the two sectors are complementary, the only way to stave off a disaster is to stop long-run growth." However, note that while our model sets the intertemporal elasticity of substitution for  electricity consumption $\sigma$ to a fixed value, we make no direct assumptions about the elasticity of substitution between different energy technologies. Instead, this latter elasticity is indirectly defined by the substitutability of electricity consumption between periods. To illustrate this, we parametrize and estimate our model numerically.

	Firstly, let technology 1 be coal power and technology 2 be solar power. These two technologies exemplify consistent fossil energy and intermittent renewable energy.  Furthermore,  let period $t$ represent peak hours and period $s$ represent off-peak hours. We assume that, holding prices equal, consumers prefer that approximately 60\% of their energy arrive in period $t$ and the remaining 40\% arrive in period $s$; that is, we have $\alpha_t = 0.6$ and $\alpha_s = 0.4$. Since the union of both periods makes up only a single day, our model's assumption of an exogenous and constant $\xi$ is fairly reasonable in this context. Next, we normalize all quantity units to a MWh basis. Hence, we let $\xi$ represent the percent of capacity utilized in each period; we assume coal uses 100\% of its capacity in both periods, while solar can access 100\% during peak hours and only 10\% during the off-peak. Finally, we set the cost parameters, given in  \$ per MWh, equal to LCOE\footnote{ The LCOE for a generation technology is equal to the sum of its lifetime costs divided by its lifetime energy output. \citet{Joskow2011} explains the flaws of comparing generation technologies solely on the basis of LCOE; but,  our use of this measure is unrelated to his critique. That is, he argues that the economic value an intermittent technology should also account for when it produces electricity and the prices of electricity in those periods (see Table 2 of his paper). Our model does exactly this; but, we still need to  use LCOE to parametrize the cost of developing capacity. } estimates for 2023 from the \citet{EIALCOE}; specifically, we use estimates for "Solar PV" and "Coal with 30\% CCS" from Table 1b.  In total, we have the following parameters.
